---
title: "Assignment B-1: Making a function"
author: "Jaden L. Phillips"
date: "2022-11-01"
output: github_document
---
## Tidy Submission (15 points)
Follow these steps to submit your work. Be sure to familiarize yourself with the rubric for a tidy submission below, before doing these steps.

1. Make a README file for your repository. It should be a brief document letting a visitor know what’s in this repository (at a high level) and some key things they should know about how to use the files in the repository.
2. Tag a release in your GitHub repository corresponding to your submission before the deadline.
3. Grab the URL corresponding to your tagged release, and submit that to canvas.

Loading directory with packages needed to run the code for this Assignment:
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(testthat)
library(palmerpenguins)
library(gapminder)
```

## Exercise 1: Make a Function (25 points)

In this exercise, you’ll be making a function and fortifying it. The function need not be complicated. The function need not be “serious”, but shouldn’t be nonsense.\

  **Function Ideas**

-   Did you repeat any code for a data analysis in STAT 545A? If so, consider making a function for this action.
-   Consider bundling a specific group_by() %>% summarise() workflow.
-   Write a wrapper around an existing function.
    -   For example, perhaps accepting a narrower range of inputs (like not allowing logical vectors), or providing a different output.
    -   A specific example: my rqdist() function is a wrapper around quantreg::rq(), narrowing its functionality.
    -   It’s usually better to narrow a function’s focus than to broaden, so that a function doesn’t end up doing too much.
-   Make a function extracting parts of an "lm" object that you can’t easily access (broom aside), or measuring the “difference” between a linear and quadratic fit.
-   Make a special plot that you’d want to repeat when exploring your data.
-   …

```{r}
avg_cols_group <- function(.data, y, na.rm = FALSE, ...) {
  if(!is.data.frame(.data)) {
    stop('I am so sorry, but this function only works for data frame inputs!\n',
         'You have provided an object of class: ', class(.data)[1])
  }
  .data %>%
     group_by({{y}}, {{...}}) %>% 
     summarise(across(where(is.numeric), mean, na.rm = TRUE))
}
```

## Exercise 2: Document your Function (20 points)
In the same code chunk where you made your function, document the function using roxygen2 tags. Be sure to include:

1. Title.
2. Function description: In 1-2 brief sentences, describe what the function does.
3. Document each argument with the @param tag, making sure to justify why you named the parameter as you did.
    -   (Justification for naming is not often needed, but we want to hear your reasoning.)
4. What the function returns, using the @return tag.

```{r}
#' Column Averages per Group (avg_cols_group)
#' 
#' This function first identifies a grouping variable (taken as *categorical*). Inspecting across all columns of the data frame, this function then selects all of the numeric columns and calculates the means of each of those numeric columns for each group (i.e., level of the grouping variable, corresponding to particular rows).
#'
#' @param .data the data frame or data frame extension (e.g. tibble). We opted for the function name ".data" to avoid name collision with the `data()` function in base R packages.
#' @param y     the variable to group by. This is interpreted as a categorical variable. If left unspecified, the function will still work, though will simply be computing means of all rows since no grouping parameter was specified.
#' @param na.rm a logical value (TRUE / FALSE) indicating whether NA values should be stripped before the computation proceeds. If `FALSE` (the default), the result will be `NA` if any element of `y` or another grouping variable is `NA`.
#' @param ...   the optional, additional, variable(s) to group by.
#'
#' @return Returns a data frame containing to the average value of each numeric column per level of the grouping variable (if there is one, or more). With the exception of column names as the top row of the dataframe, each subsequent row of the dataframe corresponds to a grouping 'level', and contains the means -- for that level -- of all the numeric variables in the dataset.

avg_cols_group <- function(.data, y, na.rm = FALSE, ...) {
  if(!is.data.frame(.data)) {
    stop('I am so sorry, but this function only works for data frame inputs!\n',
         'You have provided an object of class: ', class(.data)[1])
  }
  .data %>%
     group_by({{y}}, {{...}}) %>% 
     summarise(across(where(is.numeric), mean, na.rm = TRUE))
}
```

## Exercise 3: Include examples (15 points)
Demonstrate the usage of your function with a few examples. Use one or more new code chunks, describing what you’re doing.

Note: If you want to deliberately show an error, you can use error = TRUE in your code chunk option.

#### Example 1: **mtcars** dataset
In this example we are looking at the mtcars dataset from the base R `datasets` package:\

-   We are interested in learning about how the various vehicle specs compare across cars with different # of cylinders. 

-   Here, we can utilize the avg_cols_group( ) function, and select 'cyl' as the categorical variable to group by. 

-   The output returns a tibble with one row corresponding to the average values across each variable for each 4, 6, and 8 cylinders. cars.
```{r}
avg_cols_group(.data = mtcars, y = cyl, na.rm = TRUE)
```

#### Example 2: **gapminder** dataset
In this example we are looking at the gapminder dataset, which is an excerpt of the `gapminder` data on life expectancy, GDP per-capita, and population by country:

-   We are interested in comparing the various numeric variables across countries

-   Using the avg_cols_group( ) function, we can achieve this in one step:

    -   Indicate to group by the column y = `country`, which then computes the averages across all columns for rows corresponding to each country, which turns out to be across all years in the dataset.
    
    -   This returns a much tidier, smaller tibble with just **a single row corresponding to each country**, including the mean of (not relevant but computed because it is numeric:) year, life expectancy, population, and GDP per-capita across all years in the dataset.

```{r}
avg_cols_group(.data = gapminder, y = country, na.rm = TRUE)
```

#### Example 3: **penguins** dataset
By applying the avg_group_var( ) function to the penguins dataset, and indicating to the function to group by the `species`variable:

-   This returns an output tibble with 3 rows corresponding to the 3 penguin species in the original penguins dataset: Adelie, Chinstrap, and Gentoo.

-   Each row also contains the mean value of all numeric variables corresponding to penguins of that species.

The subsequent code uses this function in creating a scatter-plot of the flipper length v.s. body mass of the average sized penguins of each species (with average body mass & flipper length). This plot is not that interesting but a similar plot could be useful if more groups are involved in your dataset.

```{r}
avg_cols_group(.data = penguins, y = species, na.rm = TRUE) 

avg_cols_group(.data = penguins, y = species, na.rm = TRUE) %>%
  ggplot(aes(x = flipper_length_mm, y = body_mass_g, colour = species)) +
    geom_point() +
    theme_minimal() +
    xlim(0, 250) +
    ylim(0, 6000)
```


## Exercise 4: Test the Function (25 points)
Running examples is a good way of checking by-eye whether your function is working as expected. But, having a formal “yes or no” check is useful when you move on to other parts of your analysis.

Write formal tests for your function. You should use at least three non-redundant uses of an expect_() function from the testthat package, and they should be contained in a test_that() function (or more than one). They should all pass.

Example of non-redundant inputs:

-   Vector with no NA’s
-   Vector that has NA’s
-   Vector of a different type (if relevant)
-   Vector of length 0, like numeric(0).

Example of redundant inputs:

-   Providing a different number (unless one of these numbers have some significance, like a limit point – just tell us if that’s the case)

#### Testing function with **expect_equal( )**:

In the following code, I first create a dummy data.frame, which I name *dummy_df*.
```{r}
a <- c("red", "green", "blue", "orange", "yellow")
b <- 1:5
c <- b * 2
d <- 1
dummy_df <- data.frame(a, b, c, d)
dummy_df
```

I then test my *avg_cols_group( )* function using the **expect_equal( )** function:

-   This test is ran to confirm that the correct computations are happening behind the scenes of my function, such that it gives the same/equal results as were one to manually identify the numeric columns and then calculate their means
```{r}
test_that("Function output is equual to the exact values & format specified, as calculated by hand", {
  expect_equal(avg_cols_group(dummy_df), tibble(b = mean(1:5), c = mean(2 *(1:5)), d = mean(1)))
})
```

#### Testing function with **expect_named( )**:

This function is expected to only output columns with names corresponding to the numeric columns or the grouping variable(s). 

-   In the example of the penguins dataset, the avg_cols_group( ) function, when specified to group my species, should return only columns corresponding to the grouping variable "species", followed my the numeric variables "bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g" and "year".

- In other words, in this example, "island" & "sex" columns should not be in the output.

-   We check this using the expect_named( ) function, to see if only the column names we expect to see are outputted.
```{r}
test_that("The function outputs only the grouping column(s) plus columns containing numeric data", {
  expect_named(avg_cols_group(.data = penguins, y = species), c("species", "bill_length_mm", "bill_depth_mm", "flipper_length_mm", "body_mass_g", "year"))
})
```

#### Testing function with **expect_error( )**:

In the following code, I first create a dummy vector, which I name *dummy_vector*.

I choose to run a vector as an input in order see if my function throws an error when doing so, as my function should only accept inputs of the class data.frame in the `.data` parameter.

-   Indeed, as expected, and confirmed by this **expect_error( )** function, calling on an object of class 'vector' returns an error for the avg_cols_group( ) function. 


```{r}
dummy_vector <- c(1, 5, 4, 9, 0)
test_that("Error occurs when input object is not of the class data.frame", {
  expect_error(avg_cols_group(dummy_vector))
})
```



