---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# quick.group.calc

<!-- badges: start -->
<!-- badges: end -->

## Overview 

The goal of **quick.group.calc** is to simplify and make more efficient the code needed to create a group_by() %>% summarise() workflow:

-   Specifically, since most data operations are done on groups defined by variables, this package can help fast track to a compact, grouped and summarized output data frame. 

-   The output will have one row corresponding to each (combination of) grouping variable(s), and one column for each grouping variable as well as one column for each summary statistic specified. 

The primary `avg_cols_group()` function of this package first identifies a grouping variable (interpreted as categorical), and then, inspecting across all columns of the data frame, selects all of the numeric columns and calculates the means of each of those columns for each group (i.e., level of the grouping variable, corresponding to particular rows).


## Installation

You can install the development version of `quick.group.calc` from [GitHub](https://github.com/stat545ubc-2022/asgmt.b1.b2-jlp/tree/main/quick.group.calc) with:

```{r}
#install.packages("devtools")
library(devtools)
devtools::install_github("stat545ubc-2022/quick.group.calc-jlp")
```


## Examples (Usage Demonstration)

The following are two simple examples showing you how to use the primary **avg_cols_group( )** function of the **quick.group.calc** package:

```{r example}
library(quick.group.calc)

## basic example code
avg_cols_group(.data = mtcars, y = cyl)

avg_cols_group(.data = palmerpenguins::penguins, y = species)
```
